\documentclass{article}
\usepackage{booktabs}
\usepackage{amsmath}

\title{A Reproducible Neuro-Symbolic Revision Benchmark with OpenNARS}

\author{
  Anonymous Authors
}

\begin{document}
\maketitle

\begin{abstract}
We present a fully executable neuro-symbolic prototype that maps uncertain concept memberships into Narsese truth values and delegates inference to OpenNARS for Applications (ONA). We focus on a controlled fit/disambiguation domain and evaluate six explicit ablations under held-out noun-pair generalization. The key empirical result is that ONA-based belief revision and multi-hop chaining are both necessary: direct propagation baselines saturate around $0.59$ accuracy, revision-only reaches $\sim0.92$, and multi-hop+revision reaches $1.00$ on the current benchmark. All reported numbers are generated from runnable code in this repository with confidence intervals and paired significance tests.
\end{abstract}

\section{Scope and Claim}
This work is a controlled benchmark study, not a claim of broad Winograd-scale natural language understanding. The contribution is methodological and empirical:
\begin{enumerate}
    \item deterministic bridge from soft concept signals to Narsese truth values,
    \item executable ONA reasoning loop (including contradiction handling),
    \item ablation protocol separating direct propagation, revision, and multi-hop reasoning.
\end{enumerate}

\section{Method}
For each sentence, the bridge emits statements of the form:
\[
\langle adjective \rightarrow concept \rangle.\ \%f;c\%
\]
where $f$ and $c$ are bounded frequency/confidence values. ONA then reasons over background rules and optional conflict context.

We evaluate three ONA variants:
\begin{itemize}
    \item \textbf{Direct propagation}: $\langle large\_like \rightarrow subject\_cause \rangle$ and $\langle small\_like \rightarrow object\_cause \rangle$ only.
    \item \textbf{Revision}: direct propagation + contradictory context (e.g., $\langle large \rightarrow small\_like \rangle$).
    \item \textbf{Multi-hop + revision}: revision plus additional latent chains (e.g., $\langle compact\_like \rightarrow small\_like \rangle$).
\end{itemize}

\section{Evaluation Protocol}
We generate a deterministic synthetic benchmark with 60 noun pairs ($360$ total examples). Each pair contributes:
\begin{itemize}
    \item lexical core examples,
    \item synonym generalization examples,
    \item conflicting evidence examples,
    \item multi-hop chain examples.
\end{itemize}

Train/test split is done by noun pair (50/50), so test examples use unseen noun identifiers. Neural baselines are trained only on training rows and evaluated on held-out rows. We report:
\begin{itemize}
    \item test accuracy,
    \item 95\% bootstrap confidence intervals,
    \item paired McNemar exact tests against the strongest ONA condition.
\end{itemize}

\section{Results}
Primary held-out run (\texttt{split\_seed=0}, $n_{test}=180$):

\begin{table}[h]
\centering
\caption{Held-out test accuracy with 95\% bootstrap CI}
\begin{tabular}{lcc}
\toprule
Method & Accuracy & 95\% CI \\
\midrule
Exact lexical baseline & 0.483 & [0.411, 0.556] \\
Embedding-only bridge & 0.600 & [0.528, 0.667] \\
Neural baseline (MLP ensemble) & 0.600 & [0.528, 0.667] \\
ONA direct propagation & 0.600 & [0.528, 0.667] \\
ONA revision (with conflict context) & 0.933 & [0.894, 0.967] \\
\textbf{ONA multi-hop + revision} & \textbf{1.000} & \textbf{[1.000, 1.000]} \\
\bottomrule
\end{tabular}
\end{table}

Scenario-wise behavior:
\begin{itemize}
    \item conflicting\_evidence: only revision-enabled ONA variants solve this (\(1.00\)).
    \item multihop\_chain: revision-only ONA is insufficient; multi-hop rules are required.
\end{itemize}

McNemar exact tests vs ONA multi-hop+revision are significant for all comparisons in this run (largest \(p=4.88\times10^{-4}\)).

\subsection{Split Robustness}
Across five noun-heldout split seeds (0..4):
\begin{itemize}
    \item exact lexical baseline: mean \(0.487\), std \(0.008\)
    \item embedding-only bridge: mean \(0.589\), std \(0.010\)
    \item neural baseline (MLP ensemble): mean \(0.584\), std \(0.011\)
    \item ONA direct propagation: mean \(0.589\), std \(0.010\)
    \item ONA revision: mean \(0.922\), std \(0.010\)
    \item ONA multi-hop+revision: mean \(1.000\), std \(0.000\)
\end{itemize}

\section{External Benchmark Check}
We added an offline evaluation on the cached Winograd WSC273 set. Full-set neural anchors (offline checkpoints only) are:
\begin{itemize}
    \item SentenceTransformer replacement: \(0.498\)
    \item nearest mention: \(0.498\)
    \item GPT-2 sentence score: \(0.524\)
    \item GPT-2 medium sentence score: \(0.549\)
    \item BERT-base MLM option score: \(0.553\)
    \item RoBERTa-large MLM option score: \(0.689\)
\end{itemize}

To reduce handcrafted bridge dependence, we run a stratified 5-fold learned bridge on full WSC273 (22 features from replacement/context and offline model scores) and then map predicted probabilities into Narsese truth values for ONA inference:
\begin{itemize}
    \item learned bridge k-fold: \(0.670\)
    \item learned ONA direct k-fold: \(0.670\)
    \item learned ONA multihop k-fold: \(0.670\)
    \item learned ONA revision k-fold: \(0.656\)
\end{itemize}
Compared against the strongest full-WSC anchor (RoBERTa-large MLM, \(0.689\)), learned ONA direct is close but still lower (\(\Delta=-0.018\), McNemar \(p=0.625\)). For the strict executable causal paired subset (\(26\) examples), learned bridge + ONA reaches \(0.654\) vs \(0.538\) for static centroid/direct ONA mappings.
Calibration on full-WSC CV indicates learned bridge and learned ONA multihop are similar (Brier \(\approx0.21\)), while direct/revision ONA mappings are less calibrated.

\section{Limitations}
\begin{itemize}
    \item External benchmark integration exists, but current transfer performance is near (not above) the strongest full-WSC neural anchor.
    \item Synthetic benchmark semantic grounding is still hand-calibrated at concept level.
    \item The current neural baseline is a compact offline MLP, not a large pretrained LM.
\end{itemize}

\section{Conclusion}
The current repository now supports reproducible, non-mocked evidence that ONA-style revision and multi-hop symbolic inference provide measurable gains over lexical, embedding-only, and compact neural baselines in this controlled setting. The next step toward top-tier claims is transfer to externally accepted datasets with similarly strict no-mock protocols.

\end{document}
